Certified Knowledge Compilation \\ with Application to Verified Model Counting

### 2023-03-16 doc/mc-paper.tex
fix merge error

% Lean code:
\usepackage{listings}
\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{tacticcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green
\definecolor{attributecolor}{rgb}{0.7, 0.1, 0.1} % red
\def\lstlanguagefiles{lstlean.tex}
% set default language
\lstset{language=lean, xleftmargin=1em}
\lstset{backgroundcolor=\color{white}}

@@

\begin{definition}[Ring Evaluation Problem]
  For commutative ring $\ring$, a {\em ring labeling} associates a value $w(x) \in \dset$ with
  every variable $x \in \varset$.  We then define $w(\obar{x})$ to equal $\mulident-w(x)$.

@@

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
For any labeling $w$, ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies:
\begin{eqnarray}
\textstyle
\rep(\boolnot \phi,\; w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
\textstyle
\rep\left(\Pand_{i=1,k} \phi_i,\; w \right) &=& \prod_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:product} \\
\textstyle
\rep\left(\Por_{i=1,k} \phi_i,\; w\right) &=& \sum_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:sum}
\end{eqnarray}
\end{proposition}

@@

A CRAT file provides both a declaration of a POG, as well as an easily checkable
proof that a Boolean formula, given in conjunctive normal
form, is logically equivalent to the POG, thus enabling certified
knowledge compilation.
%% The proof system is based on extended
%% resolution~\cite{Tseitin:1983} with reverse unit propagation
%% (RUP)~\cite{goldberg,vangelder08_verifying_rup_proofs} as the core method for
%% proving that a set of clauses logically implies another clause.

>>>>

% Lean code:
\usepackage{listings}
%\definecolor{keywordcolor}{rgb}{0.7, 0.1, 0.1}   % red
\definecolor{keywordcolor}{rgb}{0.0, 0.1, 0.6}   % blue
\definecolor{tacticcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{commentcolor}{rgb}{0.4, 0.4, 0.4}   % grey
\definecolor{symbolcolor}{rgb}{0.0, 0.1, 0.6}    % blue
\definecolor{sortcolor}{rgb}{0.1, 0.5, 0.1}      % green
\definecolor{attributecolor}{rgb}{0.7, 0.1, 0.1} % red
\def\lstlanguagefiles{lstlean.tex}
% set default language
\lstset{language=lean, xleftmargin=1em}
\lstset{backgroundcolor=\color{white}}

@@

\begin{definition}[Ring Evaluation Problem]
\label{def:ring_evaluation}
  For commutative ring $\ring$, a {\em ring labeling} associates a value $w(x) \in \dset$ with
  every variable $x \in \varset$.  We then define $w(\obar{x})$ to equal $\mulident-w(x)$.

@@

Performing ring evaluation on an arbitrary Boolean formula could be intractable, but it is straightforward for a formula with partitioned operations:
\begin{proposition}
\label{prop:ring:eval}
Ring evaluation with operations $\boolnot$, $\pand$, and $\por$ satisfies the following for any labeling $w$:
\begin{eqnarray}
%% V1
\textstyle
\rep(\boolnot \phi,\; w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
\textstyle
\rep\left(\Pand_{i=1,k} \phi_i,\; w \right) &=& \prod_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:product} \\
\textstyle
\rep\left(\Por_{i=1,k} \phi_i,\; w\right) &=& \sum_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:sum}
%% V2
%\rep(\boolnot \phi, w) &=& \mulident - \rep(\phi, w) \label{eqn:ring:negation} \\
%\rep\left(\phi_1, w \pand \dots \pand \phi_k, w \right) &=& \prod_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:product} \\
%\rep\left(\phi_1, w \por \dots \por \phi_k, w \right) &=& \sum_{i=1..k} \rep(\phi_i, w) \label{eqn:ring:sum}
%\rep\left(\Pand_{i=1,k} \phi_i, w \right) &=& \prod_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:product} \\
%\rep\left(\Por_{i=1,k} \phi_i, w\right) &=& \sum_{i=1,k} \rep(\phi_i, w) \label{eqn:ring:sum}
\end{eqnarray}
\end{proposition}

@@

A CRAT file provides both a declaration of a POG, as well as an easily checkable
proof that a Boolean formula, given in conjunctive normal
form, is logically equivalent to the POG, thus enabling certified
knowledge compilation.
%% The proof system is based on extended
%% resolution~\cite{Tseitin:1983} with reverse unit propagation
%% (RUP)~\cite{goldberg,vangelder08_verifying_rup_proofs} as the core method for
%% proving that a set of clauses logically implies another clause.
%% Alternate version
%% knowledge compilation.  The proof system is based on extended
%% resolution~\cite{Tseitin:1983} with reverse unit propagation
%% (RUP)~\cite{goldberg,vangelder08_verifying_rup_proofs} as the core method for
%% proving that a set of clauses logically implies another clause.
%%
%% As an example, consider the formula $\varphi = (\obar{a} \lor b) \land (\obar{b} \lor c \lor \obar{e}) \land (\obar c \lor d)$
%% and a RUP step to derive the target clause $(\obar{a} \lor d \lor \obar{e})$ from the three clauses from $\varphi$.
%% A RUP proof would take the following form.  We start with the assignment that falsifies the target clause. This assignment
%% is extended by clauses in $\varphi$ that become unit and eventually the empty clause. In the illustration below, the
%% involved clauses (antecedents) are shown in the order that they became unit / falsified.
%%
%% \begin{center}
%%   \begin{tabular}{lcccc}
%%          & \makebox[15mm]{Target}    & \makebox[10mm]{} & \makebox[10mm]{Antecedents} & \makebox[10mm]{} \\
%%   Clause & $\obar{a} \lor d \lor \obar{e}$ & $\obar{c} \lor d$  & $\obar{e} \lor \obar{b} \lor c$   & $\obar{a} \lor b$ \\
%%   \midrule
%%   Units  & $a$, $\obar{d}$, $e$      &  $\obar{c}$             & $\obar{b}$                         & $\nil$ \\
%%   \end{tabular}
%% \end{center}
%%
%% RUP is an alternative formulation of resolution.  For target clause
%% $C$, it can be seen that applying resolution operations to the
%% antecedent clauses from right to left will derive a clause $C'$ such
%% that $C' \subseteq C$.  By {\em subsumption} %~\cite{philipp:lpar:2017},
%% we then have $C' \rightarrow C$.  Compared to listing each resolution
%% operation as a separate step, using RUP as the basic proof step makes
%% the proofs more compact.
%%
%% As is shown in Figure~\ref{fig:chain}
%% two programs are involved in producing a certified compilation of a formula:
%% \begin{itemize}
%% \item The {\em proof generator} produces a CRAT file that both defines the POG and gives a proof that the POG is logically equivalent to the input formula.
%% As the figure indicates, the generator starts with the output of another knowledge compiler, such as one generating a d-DNNF representation of the formula.
%% \item The {\em proof checker} ensures that all of the proof conditions are satisfied.
%% \end{itemize}
%% >>>>>>> 9c1793da858d00e5d4022d6d33c009848ba2083b
